### 2024-06-24 basic-statistics.qmd
20240624 - free power tools

Knowing any three of the following, you can calculate the fourth: statistical power, effect size, sample size, and alpha level.

@@

>>>>

Knowing any three of the following, you can calculate the fourth: statistical power, effect size, sample size, and alpha level.
Below is `R` code for calculating power for each of various statistical tests (i.e., a *power analysis*).
For free point-and-click software for calculating statistical power, see G\*Power: <https://www.psychologie.hhu.de/arbeitsgruppen/allgemeine-psychologie-und-arbeitspsychologie/gpower.html>

@@

#### Multilevel Modeling {#sec-statisticalPowerMultilevelModeling}

Power analysis for multilevel modeling approaches is more complicated than it is for other statistical analyses, such as correlation, multiple regression, *t*-tests, ANOVA, etc.

There are free web applications for calculating power in multilevel modeling:

- <https://aguinis.shinyapps.io/ml_power/>
- <https://koumurayama.shinyapps.io/tmethod_mlm/>
- <https://webpower.psychstat.org/wiki/models/index>

#### Path Analysis, Factor Analysis, and Structural Equation Modeling {#sec-statisticalPowerSEM}

Power analysis for latent variable modeling approaches like SEM is more complicated than it is for other statistical analyses, such as correlation, multiple regression, *t*-tests, ANOVA, etc.

I provide an example of power analysis in SEM using Monte Carlo simulation in R here: <https://isaactpetersen.github.io/Principles-Psychological-Assessment/sem.html#monteCarloPowerAnalysis> [@PetersenPrinciplesPsychAssessment].

There are also free web applications for calculating power in SEM:

- <https://sjak.shinyapps.io/power4SEM/>
- <https://sempower.shinyapps.io/sempower/>
- <https://yilinandrewang.shinyapps.io/pwrSEM/>
- <https://webpower.psychstat.org/wiki/models/index>

#### Mediation and Moderation

There are free tools for calculating power for tests of mediation and moderation:

- <https://www.causalevaluation.org/power-analysis.html> (web application: <https://powerupr.shinyapps.io/index/>)
- <https://webpower.psychstat.org/wiki/models/index>
